Analysing the effect of land-use type on the proportion of scouts

### 2021-11-17 analysis/landuse_effect_prop_scouts.Rmd
created new landuse plots for the PLS analysis. Awaiting a land-use conversion document from Elli, once I have this I can colour the map with the significant land-use catagories and thus show visually what changes in the environment as the proportion of scouts increases

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(size = 42)
    )
)

@@

```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}
library(sf)
library(sp)
library(ggspatial)

# make fit map with agrirural landuses
agrirural_plotdata$site <- rural_lu$site
agrirural_plotdata$prop_scouts <- rural_lu$p
ar_sortedpropscouts <- arrange(agrirural_plotdata, prop_scouts)
ar_marked <- ar_sortedpropscouts[
  c(1, nrow(ar_sortedpropscouts) / 2, nrow(ar_sortedpropscouts)),
]

# make model fit graph of pc1 against prop scouts and mark the sites with the
# lowest, middle and highest proportion of scouts
ar_fit <- ggplot(
  data = ar_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
) +
  geom_point() +
  geom_point(
    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
  ) +
  geom_text(
    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
    nudge_x = c(0, 0), nudge_y = c(-0.05, -0.05), size = 8
  ) +
  geom_line(data = rural_fitdata, aes(x = Comp_.1, y = fit)) +
  geom_ribbon(
    data = rural_fitdata,
    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  ) +
  theme(legend.position = "none") +
  xlab("PC 1") +
  ylab("Estimated proportion of scouts")

# make the maps for each of the sites selected in the previous step
site_shape_data <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/agri-rural/",
    ar_marked[i, "site"], "/", ar_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  site_shape_data[[ar_marked[i, "site"]]] <- df
}

site_shape_data$STU$descriptio

ggplot() +
  geom_sf(
    data = site_shape_data$STU
  )
```

```{r}

>>>>

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
# theme_set(
#   theme_classic() +
#     theme(
#       text = element_text(size = rel(42))
#     )
# )

@@

Here I create figures which show the relationship between PC1 and the proportion of scouts as well as land-use maps of sites long the gradient which are coloured by the land-use types identified as most singificant in the PLS (PC1).

### Agri-rural
```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}
library(sf)
library(sp)
library(ggspatial)

# make fit map with agrirural landuses
agrirural_plotdata$site <- rural_lu$site
agrirural_plotdata$prop_scouts <- rural_lu$p
ar_sortedpropscouts <- arrange(agrirural_plotdata, prop_scouts)

# select sites to draw attention to, those sitting on the trend
ar_marked <- ar_sortedpropscouts %>%
  filter(site %in% c("MEL", "SOM", "BEL"))

# make model fit graph of pc1 against prop scouts and mark the sites
ar_fit <- ggplot(
  data = ar_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
) +
  geom_point() +
  geom_line(data = rural_fitdata, aes(x = Comp_.1, y = fit)) +
  geom_ribbon(
    data = rural_fitdata,
    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  ) +
  geom_point(
    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
  ) +
  geom_text(
    data = ar_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(3)
  ) +
  theme_classic() +
  theme(legend.position = "none") +
  xlab("PC 1") +
  ylab("Estimated proportion of scouts")

# make the maps for each of the sites selected in the previous step
site_shape_data <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/agri-rural/",
    ar_marked[i, "site"], "/", ar_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  site_shape_data[[ar_marked[i, "site"]]] <- df
  site_plots[[ar_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df,
      aes(fill = descriptio)
    ) +
    theme_void()
}

# make loadings plot
argi.rural.loadings <- tibble(
  "Landuse" = names(agrirural_loadings$loadings),
  "Correlation" = agrirural_loadings$loadings
) %>%
  mutate(
    significant = ifelse(
      Correlation > agrirural_loadings$cutoff,
      Landuse,
      ifelse(
        Correlation < -agrirural_loadings$cutoff,
        Landuse,
        "Other"
      )
    )
  ) %>%
  mutate(
    Landuse = ifelse(
      Landuse == "Non.agricultural.managed.green.space",
      "NAMGS",
      ifelse(
        Landuse == "Non.agricultural.unmanaged.green.space",
        "NAUMGS",
        Landuse
      )
    )
  )

loading_plot <- ggplot(argi.rural.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(3)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.25,
    nudge_y = c(0, 0), nudge_x = c(0.5, 0.5), size = rel(3)
  ) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None"
  ) +
  scale_colour_manual(
    values = c(
      "arable" = "red3", "built.up.area" = "orange3",
      "Non.agricultural.unmanaged.green.space" = "seagreen3",
      "water" = "cornflowerblue", other = "grey"
    )
  ) +
  ylim(-0.6, 0.6)

loading_plot

# combine plots into facet

# first the maps in a column of three
sitelumaps <- plot_grid(
  site_plots[[3]] + theme(legend.position = "none"),
  site_plots[[2]] + theme(legend.position = "none"),
  site_plots[[1]] + theme(legend.position = "none"),
  ncol = 1,
  align = "v",
  labels = c(
    names(site_plots)[3], names(site_plots)[2], names(site_plots)[1]
  )
)

# now the model fit and pls results in a plot of two
mfitpls <- plot_grid(
  ar_fit + theme(plot.margin = unit(c(0., 0., 1., 0.), "cm")),
  loading_plot,
  ncol = 1,
  align = "v",
  labels = c(
    "A", "B"
  )
)

# all together

comb <- plot_grid(
  mfitpls, sitelumaps,
  nrow = 1, scale = 0.9, align = "h"
)

comb
```

### Urban

```{r, agrirural_landuse_maps, message = FALSE, warning = FALSE}

# make fit map with agrirural landuses
urban_plotdata$site <- urban_lu$site
urban_plotdata$prop_scouts <- urban_lu$p
u_sortedpropscouts <- arrange(urban_plotdata, prop_scouts)
u_marked <- u_sortedpropscouts %>%
  filter(site %in% c("HOR", "SWP", "CAD"))

# make model fit graph of pc1 against prop scouts and mark the sites with the
# lowest, middle and highest proportion of scouts
u_fit <- ggplot(
  data = u_sortedpropscouts, aes(x = Comp_.1, y = prop_scouts)
) +
  geom_point() +
  geom_point(
    data = u_marked, aes(x = Comp_.1, y = prop_scouts, colour = "red")
  ) +
  geom_text(
    data = u_marked, aes(x = Comp_.1, y = prop_scouts, label = site),
    nudge_x = c(0, 0), nudge_y = c(-0.05, 0.05), size = rel(3)
  ) +
  geom_line(data = urban_fitdata, aes(x = Comp_.1, y = fit)) +
  geom_ribbon(
    data = urban_fitdata,
    aes(x = Comp_.1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  ) +
  theme_classic() +
  theme(legend.position = "none") +
  xlab("PC 1") +
  ylab("Estimated proportion of scouts")

# make the maps for each of the sites selected in the previous step
site_shape_data <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/urban/",
    u_marked[i, "site"], "/", u_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  site_shape_data[[u_marked[i, "site"]]] <- df
  site_plots[[u_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df,
      aes(fill = descriptio)
    ) +
    theme_void()
}

# make loadings plot
urban.loadings <- tibble(
  "Landuse" = names(urban_loadings$loadings),
  "Correlation" = urban_loadings$loadings
) %>%
  mutate(
    significant = ifelse(
      Correlation > urban_loadings$cutoff,
      Landuse,
      ifelse(
        Correlation < -urban_loadings$cutoff,
        Landuse,
        "Other"
      )
    )
  )

loading_plot <- ggplot(urban.loadings, aes(x = Landuse, y = Correlation)) +
  geom_point(aes(colour = significant), size = rel(3)) +
  geom_segment(aes(
    x = Landuse, xend = Landuse, y = 0, yend = Correlation, colour = significant
  )) +
  geom_text(
    aes(x = Landuse, y = Correlation, label = Landuse),
    hjust = 0.25,
    nudge_y = c(-0, -0), nudge_x = c(0.5, 0.5), size = rel(3)
  ) +
  geom_rect(
    aes(
      ymin = -agrirural_loadings$cutoff,
      ymax = agrirural_loadings$cutoff,
      xmin = 0, xmax = Inf
    ),
    alpha = 0.01
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None"
  ) +
  scale_colour_manual(
    values = c(
      "continuous..central" = "red3", "dense.residential" = "orange3",
      "sparse.residential" = "purple3", "amenity.grassland" = "seagreen3",
      "water" = "cornflowerblue", other = "grey"
    )
  ) +
  scale_y_continuous(breaks = seq(-.9, .9, by = 0.4), limits = c(-1, 1))

loading_plot

# combine plots into facet

# first the maps in a column of three
sitelumaps <- plot_grid(
  site_plots[[3]] + theme(legend.position = "none"),
  site_plots[[2]] + theme(legend.position = "none"),
  site_plots[[1]] + theme(legend.position = "none"),
  ncol = 1,
  align = "v",
  labels = c(
    names(site_plots)[3], names(site_plots)[2], names(site_plots)[1]
  )
)

# now the model fit and pls results in a plot of two
mfitpls <- plot_grid(
  u_fit + theme(plot.margin = unit(c(0., 0., 1., 0.), "cm")),
  loading_plot,
  ncol = 1,
  align = "v",
  labels = c(
    "A", "B"
  )
)

# all together

comb <- plot_grid(
  mfitpls, sitelumaps,
  nrow = 1, scale = 0.9, align = "h"
)

comb
```

# Loadings
```{r}
